---
title: "Estimators"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4,
  dpi = 96
)
```

```{r setup}
library(coresynth)
```

coresynth exposes six estimators through `scm_fit()` plus an experimental-design
variant (`scm_design()`). This article covers the method-specific options for
each. For the basics of the shared interface, see
[Get started](../articles/coresynth.html). For inference, see
[Inference](inference.html); for multi-period adoption, see
[Staggered adoption](staggered.html).

We use one synthetic panel throughout, with two auxiliary covariates so we can
demonstrate predictor- and covariate-based matching.

```{r dgp}
set.seed(1)
N <- 12; TT <- 20; T_pre <- 12
f   <- cumsum(rnorm(TT, 0, 0.5))
lam <- rnorm(N, 1, 0.3)
dat <- expand.grid(time = seq_len(TT), id = paste0("u", seq_len(N)))
dat$y      <- as.vector(outer(f, lam)) + rnorm(nrow(dat), 0, 0.3)
dat$income <- as.vector(outer(f * 0.5, lam)) + rnorm(nrow(dat), 0, 0.3)  # predictor
dat$x      <- rnorm(nrow(dat))                                           # time-varying cov
dat$d      <- as.integer(dat$id == "u1" & dat$time > T_pre)
dat$y[dat$d == 1] <- dat$y[dat$d == 1] + 2.0   # true ATT = 2.0
```

## SCM — Synthetic Control Method

*Abadie, Diamond & Hainmueller (2010).* Donor weights on the simplex match the
treated unit's pre-treatment trajectory.

### Predictors via `pred()`

By default (`predictors = NULL`) all pre-treatment outcomes form the predictor
matrix. To match on covariates and specific outcome lags instead, build a list
of `pred(vars, times, op)` specs:

```{r scm-pred}
fit_scm <- scm_fit(
  y ~ d | id + time, data = dat, method = "scm",
  predictors = list(
    pred("income", 1:T_pre),          # mean income over the pre-period
    pred("y", T_pre),                  # outcome in the last pre-period
    pred("y", 1:4, op = "mean")        # outcome averaged over early pre-period
  )
)
fit_scm$estimate
```

### Out-of-sample V selection, donor filtering, penalisation

```{r scm-options}
# Abadie (2021) S.3.2: split the pre-period to choose V out of sample
fit_oos <- scm_fit(y ~ d | id + time, data = dat, method = "scm",
                   v_selection = "oos")

# Abadie (2021) S.4: drop poorly-fitting donors by individual MSPE ratio
fit_filt <- scm_fit(y ~ d | id + time, data = dat, method = "scm",
                    donor_mspe_threshold = 5)

# Abadie & L'Hour (2021): penalised SCM (auto-selects the penalty out of sample)
fit_pen <- scm_fit(y ~ d | id + time, data = dat, method = "scm",
                   lambda_pen = "auto")

c(oos = fit_oos$estimate, filtered = fit_filt$estimate, penalised = fit_pen$estimate)
```

### Augmented SCM

`augment_scm()` applies the ridge bias-correction of Ben-Michael, Feller & Rothstein (2021):

```{r scm-aug}
aug <- augment_scm(fit_scm)
c(scm = aug$att_scm, augmented = aug$att_aug)
```

## SDID — Synthetic Difference-in-Differences

*Arkhangelsky et al. (2021).* Combines unit and time weights with a DiD-style
double differencing. Time-varying covariates are partialled out first
(Clarke et al. 2023):

```{r sdid}
fit_sdid <- scm_fit(y ~ d | id + time, data = dat, method = "sdid",
                    covariates = "x")
fit_sdid$estimate
```

## GSC — Generalised Synthetic Control

*Xu (2017).* Interactive fixed effects estimated by SVD; `r` sets the number of
latent factors. With covariates the full EM algorithm is used (E-step SVD,
M-step ridge OLS), and `fit$beta` holds the covariate coefficients.

```{r gsc}
fit_gsc <- scm_fit(y ~ d | id + time, data = dat, method = "gsc",
                   r = 2, covariates = "x")
c(estimate = fit_gsc$estimate, beta = unname(fit_gsc$beta))
```

`r` sensitivity is easy to sweep:

```{r gsc-r}
sapply(1:3, function(rr)
  scm_fit(y ~ d | id + time, data = dat, method = "gsc", r = rr)$estimate)
```

## MC — Matrix Completion

*Athey et al. (2021).* Nuclear-norm-regularised completion (Soft-Impute). The
penalty `lambda` defaults to `0.01 * sigma_max(Y)`; pass a number to override.

```{r mc}
fit_mc_auto <- scm_fit(y ~ d | id + time, data = dat, method = "mc")
fit_mc_man  <- scm_fit(y ~ d | id + time, data = dat, method = "mc", lambda = 1.0)
c(auto = fit_mc_auto$estimate, manual = fit_mc_man$estimate)
```

> On small donor pools MC's nuclear-norm shrinkage often pushes the ATT above
> the truth — a known property (Mazumder et al. 2010), not an implementation bug.

## TASC — Time-Aware Synthetic Control

*Rho et al. (2026).* A state-space model fitted by Kalman EM. `r` sets the latent
state dimension, `em_iter` the EM iterations, and `fix_A = TRUE` constrains the
transition matrix to be constant.

```{r tasc}
fit_tasc      <- scm_fit(y ~ d | id + time, data = dat, method = "tasc",
                         r = 2, em_iter = 10)
fit_tasc_fixA <- scm_fit(y ~ d | id + time, data = dat, method = "tasc",
                         r = 2, em_iter = 10, fix_A = TRUE)
c(free_A = fit_tasc$estimate, fixed_A = fit_tasc_fixA$estimate)
```

## SI — Synthetic Interventions

*Agarwal et al. (2025).* SI-PCR uses a truncated SVD of the donor pre-period.
`k` sets the rank (default `floor(sqrt(min(T_pre, N_co)))`).

```{r si}
fit_si <- scm_fit(y ~ d | id + time, data = dat, method = "si")
fit_si$estimate
```

### Multi-arm SI

SI uniquely supports multiple treatment arms (`K > 1`). When the treatment
column takes values `0, 1, ..., K`, `scm_fit()` automatically routes to the
multi-arm path: the control-arm SVD basis is shared across all arms (the unit
factors are arm-invariant). Use `panel_to_tensor()` to prepare/inspect the
tensor structure.

```{r si-multi}
set.seed(7)
Nc <- 10; T2 <- 16; Tp <- 10
U <- matrix(rnorm(T2 * 2), T2, 2)
V <- matrix(rnorm((Nc + 4) * 2), Nc + 4, 2)
Lam <- rbind(c(1, 1), c(1.4, .6), c(.6, 1.4))   # arm 0, 1, 2
arm_unit <- c(rep(0L, Nc), 1L, 1L, 2L, 2L)
g <- expand.grid(time = seq_len(T2), id = seq_len(Nc + 4))
g$arm <- arm_unit[g$id]
g$d   <- ifelse(g$arm == 0L, 0L, ifelse(g$time > Tp, g$arm, 0L))
g$y   <- mapply(function(t, i, a)
  sum(U[t, ] * Lam[a + 1L, ] * V[i, ]) + 50 + rnorm(1, sd = .3),
  g$time, g$id, g$arm)

fit_multi <- scm_fit(y ~ d | id + time, data = g, method = "si")
fit_multi$arm_estimates       # per-arm ATTs
fit_multi$estimate            # weighted aggregate
```

## SCM-Design — experimental design

*Abadie & Zhao (2026).* `scm_design()` has a different, design-oriented
interface (it selects which units to expose to a planned intervention). It
offers `"base"`, `"weakly_targeted"`, and `"unit_level"` variants with
blank-period permutation tests and split-conformal CIs.

```{r scm-design}
des <- scm_design(
  data = dat, outcome = "y", unit = "id", time = "time",
  T0 = T_pre, design = "base"
)
des
```
