---
title: "Inference"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4,
  dpi = 96
)
```

```{r setup}
library(coresynth)
library(broom)
```

coresynth ships method-appropriate inference for each estimator, plus a
cross-cutting **conformal** procedure. This article surveys them on a single
synthetic panel.

```{r dgp}
set.seed(123)
N <- 12; TT <- 20; T_pre <- 12
f   <- cumsum(rnorm(TT, 0, 0.5))
lam <- rnorm(N, 1, 0.3)
dat <- expand.grid(time = seq_len(TT), id = paste0("u", seq_len(N)))
dat$y <- as.vector(outer(f, lam)) + rnorm(nrow(dat), 0, 0.3)
dat$d <- as.integer(dat$id == "u1" & dat$time > T_pre)
dat$y[dat$d == 1] <- dat$y[dat$d == 1] + 2.0   # true ATT = 2.0

fit_scm  <- scm_fit(y ~ d | id + time, data = dat, method = "scm")
fit_sdid <- scm_fit(y ~ d | id + time, data = dat, method = "sdid")
fit_gsc  <- scm_fit(y ~ d | id + time, data = dat, method = "gsc", r = 2)
fit_si   <- scm_fit(y ~ d | id + time, data = dat, method = "si")
```

## SCM — MSPE ratio permutation test

*Abadie et al. (2010).* `mspe_ratio_pval()` compares the treated unit's
post/pre MSPE ratio against placebo ratios from every donor. It returns a plain
list.

```{r scm-mspe}
mspe <- mspe_ratio_pval(fit_scm, alternative = "two.sided")
c(p_value = mspe$p_value, ratio_obs = mspe$ratio_obs)
```

The returned `scm_placebo` object has a `plot()` method that reproduces the
two companion figures of the paper. `type = "gaps"` overlays the treated
unit's gap path (treated minus synthetic control) on the placebo gap paths
obtained by reassigning the intervention to each donor (ADH 2010, Figure 4):

```{r scm-placebo-gaps}
plot(mspe, type = "gaps")
```

Placebo units whose synthetic control fits poorly before treatment carry no
information about the rarity of a large post-treatment gap; ADH exclude them
with a relative pre-treatment MSPE cutoff (20x/5x/2x in their Figures 5–7),
available here as `mspe_prune`:

```{r scm-placebo-gaps-pruned}
plot(mspe, type = "gaps", mspe_prune = 2)
```

`type = "ratios"` shows the post/pre-treatment MSPE ratio of every unit (ADH
2010, Figure 8) — the statistic behind the permutation p-value, which needs no
pruning cutoff by construction:

```{r scm-placebo-ratios}
plot(mspe, type = "ratios")
```

Series colors, legend text, and the reference lines are customizable via
`colors`, `labels` (e.g. `labels = c(Treated = "California")`), `vline`, and
`hline` — see `?plot.scm_placebo`.

## SDID — four inference methods

*Arkhangelsky et al. (2021); Clarke et al. (2023).* `sdid_inference()` supports
`"placebo"`, `"bootstrap"`, `"jackknife"`, and `"jackknife_global"`. The result
is `broom`-friendly.

```{r sdid-inf}
inf_plac <- sdid_inference(fit_sdid, method = "placebo")
inf_boot <- sdid_inference(fit_sdid, method = "bootstrap", n_boot = 100, seed = 1)

tidy(inf_plac)
tidy(inf_boot)
```

`tidy()` returns a one-row data frame with `estimate`, `std.error`, `p.value`,
and a CI — ready to `rbind()` into a results table. `glance()` gives a compact
summary:

```{r sdid-glance}
glance(inf_boot)
```

## GSC — parametric and non-parametric

*Xu (2017).* `gsc_boot()` is a parametric bootstrap under H0 (sharp fits only);
`gsc_inference()` offers non-parametric bootstrap / jackknife.

```{r gsc-inf}
gb <- gsc_boot(fit_gsc, B = 100, alpha = 0.05)
c(ci_lower = gb$ci_lower, ci_upper = gb$ci_upper, p_value = gb$p_value)

tidy(gsc_inference(fit_gsc, method = "jackknife"))
```

## SI — bootstrap / jackknife

*Agarwal et al. (2025).* `si_inference()` mirrors the GSC non-parametric API and
also handles staggered and multi-arm fits.

```{r si-inf}
tidy(si_inference(fit_si, method = "bootstrap", n_boot = 100, seed = 1))
```

## Conformal inference (any sharp fit)

*Chernozhukov, Wüthrich & Zhu (2021).* `conformal_inference()` works across
`scm` / `sdid` / `gsc` / `mc` / `si` sharp fits. Under H0: τ = τ0 it re-imputes
the treated post-period as `Y1 - τ0`, **re-estimates the counterfactual on all
T periods**, and inverts a moving-block permutation test of the residuals to get
a p-value and a confidence interval.

```{r conformal}
conf <- conformal_inference(fit_scm, tau0 = 0, level = 0.95)
tidy(conf)
```

The p-value tests the sharp null τ0 = 0; the CI is obtained by test inversion
over a grid. Because the permutation uses |Π| = T cyclic shifts, the smallest
attainable p-value is 1/T.

## Assembling a results table

Since every inference object tidies to the same columns, comparing methods is a
single `rbind()`:

```{r table}
do.call(rbind, list(
  cbind(method = "sdid (placebo)",   tidy(inf_plac)[c("estimate","p.value","conf.low","conf.high")]),
  cbind(method = "sdid (bootstrap)", tidy(inf_boot)[c("estimate","p.value","conf.low","conf.high")]),
  cbind(method = "scm (conformal)",  tidy(conf)[c("estimate","p.value","conf.low","conf.high")])
))
```
