---
title: "Staggered adoption"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4,
  dpi = 96
)
```

```{r setup}
library(coresynth)
```

When units adopt treatment at different times, coresynth fits each adoption
**cohort** separately and aggregates the cohort ATTs with weights proportional
to `N_treated × T_post` (Clarke et al. 2023):

$$\hat\tau = \frac{\sum_g N_{tr,g}\, T_{post,g}\, \hat\tau_g}{\sum_g N_{tr,g}\, T_{post,g}}.$$

All six estimators support this; staggered timing is **detected automatically**
from the treatment column — no extra flag is needed.

## A staggered panel

Here `u1` is treated from period 11 and `u2` from period 16.

```{r dgp}
set.seed(42)
N <- 12; TT <- 20
f   <- cumsum(rnorm(TT, 0, 0.5))
lam <- rnorm(N, 1, 0.3)
dat <- expand.grid(time = seq_len(TT), id = paste0("u", seq_len(N)))
dat$y <- as.vector(outer(f, lam)) + rnorm(nrow(dat), 0, 0.3)

dat$d <- 0L
dat$d[dat$id == "u1" & dat$time > 10] <- 1L
dat$d[dat$id == "u2" & dat$time > 15] <- 1L
dat$y[dat$d == 1] <- dat$y[dat$d == 1] + 2.0   # true ATT = 2.0
```

## Fitting

```{r fit}
fit <- scm_fit(y ~ d | id + time, data = dat, method = "sdid")
fit$estimate          # aggregate ATT
fit$staggered         # TRUE
```

Per-cohort detail is in `cohort_estimates`:

```{r cohorts}
fit$cohort_estimates
```

## Choosing the control group

`control_group` controls which units serve as donors for each cohort:

- `"clean"` (default) — never-treated units **plus** not-yet-treated units
  (those adopting later than the current cohort).
- `"never_treated"` — never-treated units only.

```{r control-group}
fit_clean <- scm_fit(y ~ d | id + time, data = dat, method = "sdid",
                     control_group = "clean")
fit_nt    <- scm_fit(y ~ d | id + time, data = dat, method = "sdid",
                     control_group = "never_treated")
c(clean = fit_clean$estimate, never_treated = fit_nt$estimate)
```

Note the trade-off: `"clean"` gives earlier cohorts a larger donor pool, but a
future adopter's **post-adoption** outcomes enter the synthetic control of the
periods after it adopts, which attenuates the aggregate ATT when its treatment
effect is large. If the effect you estimate is sizeable relative to the
outcome noise, `"never_treated"` is the safer choice (this matches the fixed
donor-pool rule of Ben-Michael, Feller & Rothstein 2022, who exclude any unit
treated within the evaluation window).

## Across estimators

The same call works for every method:

```{r all-methods}
methods <- c("scm", "sdid", "gsc", "mc", "tasc", "si")
sapply(methods, function(m)
  scm_fit(y ~ d | id + time, data = dat, method = m)$estimate)
```

## SCM: partially pooled synthetic controls

The classic SCM was designed for a **single** treated unit, so its staggered
extension rests on Ben-Michael, Feller & Rothstein (2022, JRSS-B). Units
adopting at the same time are averaged into one cohort — fully pooling within
a cohort is justified by their theory — and each cohort gets its own synthetic
control. How the cohort fits interact is controlled by `nu`:

- `nu = NULL` (default) — each cohort is fitted independently with the classic
  V-optimised SCM (the behaviour of earlier versions).
- `nu` in `[0, 1]` — **partially pooled SCM**: all cohort weight vectors are
  chosen jointly, trading off per-cohort pre-treatment fit (`nu = 0`) against
  the fit of the *average* placebo gap across cohorts (`nu = 1`), which
  anchors the aggregate ATT.
- `nu = "auto"` — the paper's heuristic based on how well separate synthetic
  controls already balance the average.

`fixedeff = TRUE` additionally demeans every unit by its own pre-treatment
mean within each cohort (an intercept shift), which the paper recommends when
outcome levels differ across units.

```{r scm-pooled}
fit_pp <- scm_fit(y ~ d | id + time, data = dat, method = "scm",
                  nu = "auto", fixedeff = TRUE)
fit_pp$estimate
fit_pp$pooling[c("nu", "q_sep", "q_pool", "q_sep_separate", "q_pool_separate")]
```

`q_pool` (the pooled imbalance) improves relative to the separate solution at
the cost of a small increase in `q_sep` (the average per-cohort imbalance).

## Inference under staggered adoption

`sdid_inference()`, `gsc_inference()`, and `si_inference()` all extend to
staggered fits. `jackknife_global` is the staggered-specific variant that
removes each unique control unit across **all** cohorts at once, correctly
capturing cross-cohort correlation.

```{r stag-inf}
library(broom)
tidy(sdid_inference(fit, method = "bootstrap", n_boot = 100, seed = 1))
tidy(sdid_inference(fit, method = "jackknife_global"))
```

Staggered **SCM** fits use `scm_inference()`, the wild bootstrap of
Ben-Michael, Feller & Rothstein (2022): donor weights stay fixed and
per-treated-unit effect contributions are perturbed with mean-zero
multipliers. With only a couple of treated units the interval is unreliable
(a warning is issued below 5).

```{r scm-inf, warning=FALSE}
tidy(scm_inference(fit_pp, n_boot = 500, seed = 1))
```

## Notes

- `plot()` and `augment()` for staggered fits operate per cohort; the aggregate
  synthetic series is not defined (`Y_synth = NULL`).
- SCM staggered adoption supports `covariates =` (partial-out) but not the
  `predictors = pred(...)` interface.
- SI additionally supports staggered **and** multi-arm simultaneously — see
  [Estimators](estimators.html#multi-arm-si).
