---
title: "Staggered adoption"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4,
  dpi = 96
)
```

```{r setup}
library(coresynth)
```

When units adopt treatment at different times, coresynth fits each adoption
**cohort** separately and aggregates the cohort ATTs with weights proportional
to `N_treated × T_post` (Clarke et al. 2023):

$$\hat\tau = \frac{\sum_g N_{tr,g}\, T_{post,g}\, \hat\tau_g}{\sum_g N_{tr,g}\, T_{post,g}}.$$

All six estimators support this; staggered timing is **detected automatically**
from the treatment column — no extra flag is needed.

## A staggered panel

Here `u1` is treated from period 11 and `u2` from period 16.

```{r dgp}
set.seed(42)
N <- 12; TT <- 20
f   <- cumsum(rnorm(TT, 0, 0.5))
lam <- rnorm(N, 1, 0.3)
dat <- expand.grid(time = seq_len(TT), id = paste0("u", seq_len(N)))
dat$y <- as.vector(outer(f, lam)) + rnorm(nrow(dat), 0, 0.3)

dat$d <- 0L
dat$d[dat$id == "u1" & dat$time > 10] <- 1L
dat$d[dat$id == "u2" & dat$time > 15] <- 1L
dat$y[dat$d == 1] <- dat$y[dat$d == 1] + 2.0   # true ATT = 2.0
```

## Fitting

```{r fit}
fit <- scm_fit(y ~ d | id + time, data = dat, method = "sdid")
fit$estimate          # aggregate ATT
fit$staggered         # TRUE
```

Per-cohort detail is in `cohort_estimates`:

```{r cohorts}
fit$cohort_estimates
```

## Choosing the control group

`control_group` controls which units serve as donors for each cohort:

- `"clean"` (default) — never-treated units **plus** not-yet-treated units
  (those adopting later than the current cohort).
- `"never_treated"` — never-treated units only.

```{r control-group}
fit_clean <- scm_fit(y ~ d | id + time, data = dat, method = "sdid",
                     control_group = "clean")
fit_nt    <- scm_fit(y ~ d | id + time, data = dat, method = "sdid",
                     control_group = "never_treated")
c(clean = fit_clean$estimate, never_treated = fit_nt$estimate)
```

## Across estimators

The same call works for every method:

```{r all-methods}
methods <- c("scm", "sdid", "gsc", "mc", "tasc", "si")
sapply(methods, function(m)
  scm_fit(y ~ d | id + time, data = dat, method = m)$estimate)
```

## Inference under staggered adoption

`sdid_inference()`, `gsc_inference()`, and `si_inference()` all extend to
staggered fits. `jackknife_global` is the staggered-specific variant that
removes each unique control unit across **all** cohorts at once, correctly
capturing cross-cohort correlation.

```{r stag-inf}
library(broom)
tidy(sdid_inference(fit, method = "bootstrap", n_boot = 100, seed = 1))
tidy(sdid_inference(fit, method = "jackknife_global"))
```

## Notes

- `plot()` and `augment()` for staggered fits operate per cohort; the aggregate
  synthetic series is not defined (`Y_synth = NULL`).
- SCM staggered adoption supports `covariates =` (partial-out) but not the
  `predictors = pred(...)` interface.
- SI additionally supports staggered **and** multi-arm simultaneously — see
  [Estimators](estimators.html#multi-arm-si).
